---
title: "chapter6.Rmd"
author: "Wenhsuan Hung"
date: "2019/12/5"
output: html_document
---

# Analysis of longitudinal data

In this week, there are two datasets, which are BPRS and RATS. First, I transform both dataset into long form. I will start from interpreting RATS dataset.

**I.RATS dataset**

```{r echo = FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(lme4)
```

```{r echo = FALSE}
BPRS <- read.table("https://raw.githubusercontent.com/KimmoVehkalahti/MABS/master/Examples/data/BPRS.txt", sep=" ", header=TRUE)
RATS <- read.table("https://raw.githubusercontent.com/KimmoVehkalahti/MABS/master/Examples/data/rats.txt",sep="", header=TRUE )

#Convert the categorical variables of both data sets to factors
#BPRS
BPRS$treatment <- factor(BPRS$treatment)
BPRS$subject <- factor(BPRS$subject)
#RATS
RATS$ID <- factor(RATS$ID)
RATS$Group <- factor(RATS$Group)

# Convert BPRS to long form(360 observations and 5 variables)
BPRSL <-  BPRS %>% gather(key = weeks, value = bprs, -treatment, -subject)
# Extract the week number
BPRSL <-  BPRSL %>% mutate(week = as.integer(substr(weeks,5,5)))

# Convert RATS data to long form, and extract time 
RATSL <- RATS %>% gather(key = WD, value = Weight, -ID, -Group) %>% mutate(Time = as.integer(substr(WD,3,4))) 
```

**Plot the RATS data **
I draw two plots to show the result of different groups. X-axis is Time, and Y-axis is Weight. Each color represents each group.

```{r echo = FALSE}
# Plot the RATSL data
ggplot(RATSL, aes(x = Time, y = Weight, group = ID)) +
  geom_line(aes(linetype = Group, color=Group)) + scale_x_continuous(name = "Time (days)", breaks = seq(0, 60, 10)) + scale_y_continuous(name = "Weight (grams)")
```

The result shows that the data of each group are distributed separately. Group 1 has the lowest weight as the time pass.


**The Golden Standardize **
Since I want to see more clearly about the tracking phenomenon, I standardize values of each observation. Each line represents each group, the X-axis is Time, and Y-axis is standardized bprs. 

```{r echo = FALSE}
RATSL
# Standardize the variable bprs
RATSL <- RATSL %>%
  group_by(Time) %>%
  mutate(stdWeight = (Weight - mean(Weight))/sd(Weight) ) %>%
  ungroup()

# Plot again with the standardized Weight
ggplot(RATSL, aes(x = Time, y = stdWeight, linetype = Group)) +
  geom_line(aes(linetype=Group,color=Group)) +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  scale_y_continuous(name = "standardized Weight")
```

**Graphs showing average value**

**1.Summary graphs -Plot the mean profiles**

Since tracking phenomenon only shows the rise and fall on the data, I couldn't analysis the data clearly, hence graphs showing average value for each group, and the variation of the observations is important. It is better to study the average instead of each specific data in a huge dataset. I draw a summary graph showing the dataset.

```{r echo = FALSE}
# Number of weeks, baseline (week 0) included
n <- RATSL$week %>% unique() %>% length()

# Summary data with mean and standard error of bprs by treatment and week 
RATSS <- RATSL %>%
  group_by(Group, Time) %>%
  summarise( mean = mean(Weight), se = sd(Weight)/sqrt(n) ) %>%
  ungroup()

# Plot the mean profiles
ggplot(RATSS, aes(x = Time, y = mean, linetype = Group, shape = Group)) +
  geom_line(aes(linetype=Group,color=Group)) +
  scale_linetype_manual(values = c(1,2,3)) +
  geom_point(size=3, aes(color = factor(Group))) +
  scale_shape_manual(values = c(1,2,3)) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se, linetype="1"), width=0.3) +
  theme(legend.position = c(0.8,0.5)) +
  scale_y_continuous(name = "mean(b) +/- se(bprs)")
```

The summary graph shows the mean of each group in different time. This is clearer than the first two plots, since mean is better to interpret a big dataset (it shows the overall situation clearer). According to the plot, the mean of each group increases as the time pass.

**2.Find the Outlier**

Outliers might bias for the result, hence, in the next step, I try to find out the outliers and remove them from the dataset.**

```{r echo = FALSE}

# Create a summary data by treatment and subject with mean as the summary variable (ignoring baseline week 0).
RATSL8S <- RATSL %>%
  filter(Time > 0) %>%
  group_by(Group, ID) %>%
  summarise( mean=mean(Weight) ) %>%
  ungroup()

# Draw a boxplot of the mean versus treatment
ggplot(RATSL8S, aes(x = Group, y = mean)) +
  geom_boxplot(aes(linetype=Group,color=Group)) +
  stat_summary(fun.y = "mean", geom = "point", shape=23, size=4, fill = "white") +
  scale_y_continuous(name = "mean(Weight), weeks 1-8")

# Create a new data by filtering the outlier and adjust the ggplot code the draw the plot again with the new data
RATSL8S1 <- RATSL8S %>% filter(mean < 550)
RATSL8S2 <- RATSL8S1 %>% filter(mean > 250)

RATSL8S
```

According to the boxplot, I find that all the three groups have outlier, hence I filter them out from the dataset, to make the result more unbiased. On the other hand, since group 2 and group 3 have similar mean, and the mean of group 1 is very different from the other two group, I decide to do T test, to see whether the two groups (group 2, group 3) are similar or not.Therefore, I filter group 1 from the below t test.  

** T-test and Anova**

In the last step, I would like to apply t-test to understand the difference between group 2 and 3 (since they have similar mean). I filter group 2, 3 out in the beginning. After filtering group 2, 3, I perform a t test, and add the baseline from the original data to the summary data to do ANOVA, since baseline measurement of the outcome variable are often correlated with the chosen summary measure in longitudinal study.

```{r echo = FALSE}

RATSL8S2$Group <- as.numeric(as.character(RATSL8S2$Group)) 
#filter out group 2 and 3
RATSL8S3 <- filter(RATSL8S2, Group > 1)

# Perform a two-sample t-test
t.test(mean ~ Group, data = RATSL8S3, var.equal = TRUE)


RATS$Group <- as.numeric(as.character(RATS$Group)) 
#filter group 2, 3 from old data
RATS_new <- filter(RATS, Group > 1)

#remove ID=12 in RATS_new
RATS_new <- RATS_new[-4,]
RATS_new

# Add the baseline from the original data as a new variable to the summary data
RATSL8S4 <- RATSL8S3 %>% mutate(baseline = RATS_new$WD1)

# Fit the linear model with the mean as the response 
fit <- lm(mean ~ baseline + Group, data = RATSL8S4)

# Compute the analysis of variance table for the fitted model with anova()
anova(fit)

```

T test shows that p value is lower than 0.05 (p-value = 0.002452), which means that the average of group2, 3 are th almost the  same, hence we can see them as the same group. According to the result of ANOVA, baseline is strongly related to the RATS values.


**II.BPRS dataset**

**Plot BPRS data**
I draw two plots to show the result of different treatments. X-axis is week, and Y-axis is bprs. Each line represents each subject.

```{r echo = FALSE}

# Plot the RATSL data
ggplot(BPRSL, aes(x = week, y = bprs, linetype = subject)) +
  geom_line(aes(linetype=subject,color=subject)) +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ treatment, labeller = label_both) +
  theme(legend.position = "none") + 
  scale_y_continuous(limits = c(min(BPRSL$bprs), max(BPRSL$bprs)))
``` 

According to the result, bprs score of almost all the men decreases over the eight weeks of the study.

**Linear model**

I fit a multiple linear regression model with bprs as response and week, treatment as explanatory variables.

```{r echo = FALSE}
# create a regression model RATS_reg
BPRS_reg <- lm( bprs~ week + treatment, data = BPRSL)

# print out a summary of the model
summary(BPRS_reg)

```

 The result shows that the p-value is low, hence we can say that the model is quite precise. Also, since standard error low, the estimation is highly correct.
 . 

**Random Intercept Model **

Since there are lots of factors which affect the result of bprs, I fit the data in random intercept model, so that each rat is different in intercept from other rats.

```{r echo = FALSE}
# Create a random intercept model
BPRS_ref <- lmer(bprs ~ week + treatment + (1 | subject), data = BPRSL, REML = FALSE)

# Print the summary of the model
summary(BPRS_ref)

```

The result of standard deviation is high, which means the data are different with each other. 


** Slippery slopes: Random Intercept and Random Slope Model **

I fit random intercept and random slope model to the BPRS data. This allows linear regression fits for each individual differ in intercept and slope. By doing so,  I can analysis the data and understand whether the difference of the data is account for each treatment, or the effect of time.

```{r echo = FALSE}

#create a random intercept and random slope model
BPRS_ref1 <- lmer(bprs ~ week + treatment + (week | subject), data = BPRSL, REML = FALSE)

# print a summary of the model
summary(BPRS_ref1)

# perform an ANOVA test on the two models
anova(BPRS_ref1, BPRS_ref)

```


**Random Intercept and Random Slope Model with interaction **

In the last step, I fit random intercept and slope model that allows group × time interaction.

```{r echo = FALSE} 

# create a random intercept and random slope model
BPRS_ref2<- lmer(bprs ~ week * treatment + (week | subject), data = BPRSL, REML = FALSE)

# print a summary of the model
summary(BPRS_ref2)

# perform an ANOVA test on the two models
anova(BPRS_ref2, BPRS_ref1)
```
